Details of specific functions

### 2021-03-02 inst/gitbook/03_specific-tutorials.Rmd
Added phylo.projections (not checked)
Not checked!

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

#### Projections: elaboration and exploration

@@

>>>>

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `phylo.projections` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

#### Projections and phylo.projections: elaboration and exploration

@@

The `phylo.projections` is the same as the `projections` metric but allows to determine the vector (`(point1, point2)`) using a tree rather than manually entering these points.
The function intakes the exact same options as the `projections` function described above at the exception of `point1` and `point2`.
Instead it takes a the argument `type` that designates the type of vector to draw from the data based on a phylogenetic tree `phy`.
The argument `type` can be a pair of any of the following inputs:

 * `"root"`: to automatically use the coordinates of the root of the tree (the first element in `phy$node.label`);
 * `"ancestor"`: to automatically use the coordinates of the elements' (i.e. any row in the matrix) most recent ancestor;
 * `"tips"`: to automatically use the coordinates from the centroid of all tips;
 * `"nodes"`: to automatically use the coordinates from the centroid of all nodes;
 * `"livings"`: to automatically use the coordinates from the centroid of all "living" tips (i.e. the tips that are the furthest away from the root);
 * `"fossils"`: to automatically use the coordinates from the centroid of all "fossil" tips and nodes (i.e. not the "living" ones);
 *  any numeric values that can be interpreted as `point1` and `point2` in `projections` (e.g. `0`, `c(0, 1.2, 3/4)`, etc.);
 *  or a user defined function that with the inputs `matrix` and `phy` and `row` (the element's ID, i.e. the row number in `matrix`).

For example, if you want to measure the projection of each element in the matrix (tips and nodes) on the axis from the root of the tree to each element's most recent ancestor, you can define the vector as `type = c("root", "ancestor")`.

```{r}
## Adding a extra row to dummy matrix (to match dummy_tree)
tree_space <- rbind(dummy_space, root = rnorm(5))
## Creating a random dummy tree (with labels matching the ones from tree_space)
dummy_tree <- rtree(6)
dummy_tree$tip.label <- rownames(tree_space)[1:6]
dummy_tree$node.label <- rownames(tree_space)[rev(7:11)]

## Measuring the disparity as the projection of each element
## on its root-ancestor vector
summary(dispRity(tree_space, metric = phylo.projections,
                             phy    = dummy_tree,
                             type   = c("root", "ancestor")))
```

Of course you can also use any other options from the projections function:

```{r}
## A user defined function that's returns the centroid of
## the first three nodes
fun.root <- function(matrix, phy, row = NULL) {
  return(colMeans(matrix[phy$node.label[1:3], ]))
}
## Measuring the unscaled rejection from the vector from the
## centroid of the three first nodes
## to the coordinates of the first tip
summary(dispRity(tree_space, metric  = phylo.projections,
                             phy     = dummy_tree,
                             measure = "distance",
                             type    = list(fun.root,
                                            tree_space[1, ])))
```
